Towards a Machine Learning approach for Aggregate Computing

### 2021-11-28 paper-2021-woa-ac-ml.tex
A little bit of more motivation about gradient example
Specify program execution and computation loop are those of AC

Consider the problem of associating
 to each node of the network
 a value
 representing the estimation of the minimum distance
 from that node to the closest \emph{source} node (see \Cref{fig:hop-count}).
%
This is indeed a computational field 
 and it is known as a \emph{gradient}~\cite{DBLP:conf/saso/AudritoCDV17}
 and there exist aggregate algorithms
 for computing it in such a way that
 the field automatically adapts 
 to changes induced by mobility
  and addition/removal of sources.
%
The gradient is a key building block 
 for building more complex self-organising behaviour~\cite{DBLP:journals/jlap/ViroliBDACP19}.

@@

\section{Aggregate Computing with Reinforcement}\label{aggregate-and-rl}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{img/algorithm-learning.pdf}
  \caption{This figure shows the main steps of RL in AC.
  Each agent starts with his copy of a Q table. 
  Then, during the simulation, they refine them Q tables changing Q values according to their experience (different colours mean different values). 
  In the end, a new $Q_{max}$ table will be created aggregating each agent table. 
  Finally, a new episode starts using this $Q_{max}$ as a starting Q table.
  }
  \label{fig:aggregate-q-learning}
\end{figure}
%As first experiments, we leverage RL to improve building blocks. 
% 
In this section, we briefly describe how RL can be injected into the computation loop,
 and what RL we used with AC.

From a local view, program execution can be seen as:
$$
f : \textit{Context} \rightarrow \textit{Output}
$$
Where $\textit{Context}$ contains:
\begin{itemize}
  \item $\sigma(name) \rightarrow \textit{data}$: a function from sensor name to the perceived value;
  \item $\phi(\textit{ID}) \rightarrow \textit{Export}$: a function from id to neighbourhoods export data.
\end{itemize}
The computational field $\theta$ contains all output produced by the system.

>>>>

Consider the problem of associating
 to each node of the network
 a value
 representing the estimation of the minimum distance
 from that node to the closest \emph{source} node (see \Cref{fig:hop-count}).
%
This is indeed a computational field 
 and it is known as a \emph{gradient}~\cite{DBLP:conf/saso/AudritoCDV17}
 and there exist aggregate algorithms
 for computing it in such a way that
 the field automatically adapts 
 to changes induced by mobility
  and addition/removal of sources.
%
This example is significant as the gradient is a key building block 
 for more complex self-organising behaviour~\cite{DBLP:journals/jlap/ViroliBDACP19}.

@@

\section{Aggregate Computing with Reinforcement}\label{aggregate-and-rl}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{img/algorithm-learning.pdf}
  \caption{This figure shows the main steps of RL in AC.
  Each agent starts with his copy of a Q table. 
  Then, during the simulation, they refine them Q tables changing Q values according to their experience (different colours mean different values). 
  In the end, a new $Q_{max}$ table will be created aggregating each agent table. 
  Finally, a new episode starts using this $Q_{max}$ as a starting Q table.
  }
  \label{fig:aggregate-q-learning}
\end{figure}
%As first experiments, we leverage RL to improve building blocks. 
% 
In this section, we briefly describe how RL can be injected into the AC computation loop,
 and what RL we used with AC.

From a local view, AC program execution can be seen as:
$$
f : \textit{Context} \rightarrow \textit{Output}
$$
Where $\textit{Context}$ contains:
\begin{itemize}
  \item $\sigma(name) \rightarrow \textit{data}$: a function from sensor name to the perceived value;
  \item $\phi(\textit{ID}) \rightarrow \textit{Export}$: a function from id to neighbourhoods export data.
\end{itemize}
The computational field $\theta$ contains all output produced by the system.
